TEMPERATURES AND METALLICITIES OF M DWARFS IN THE \apogee\ SURVEY

### 2019-08-01 paper/draft.tex
address rotational broadening bias

For the physical parameter model, we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model 5 out of 87 sources show possible rotational line broadening identified by visual inspection (as indicated by the red circles in Figure~\ref{fig:mann_validation}), while the remaining sources show no obvious broadening. We note these sources with high rotational broadening have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), we find that the scatter of labels derived for these sources is consistent with the rest of the sample and we do not remove them from the training sample.

>>>>

For the physical parameter model, we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model that 5 out of 87 sources show possible rotational line broadening identified by visual inspection (as indicated by the red circles in Figure~\ref{fig:mann_validation}), while the remaining sources show no obvious broadening. We note these broadened sources have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), and have cross-validation labels biased by an average of $+65\,$K and $-0.08\,$dex, however removing them from the training sample does not significantly change the overall scatter and bias of the model (when removed the scatters are 79\,K/0.09\,dex and biases are +5\,K/+0.01\,dex). Hence we do not remove them from the training sample.
